Automating Verification of MPC Security

### 2023-06-10 popl24/bruteforce.tex
checkpoint

\subsection{Enumerating $\runs(\prog)$}

$$
g\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) =\begin{array}[t]{l}
\elet{\stores'}{\solve{\stores}{\be}}{}\\
  \{\extend{\store}{\view{\cid}{w}}{\etrue} \mid \store \in \stores' \} \cup
  \{\extend{\store}{\view{\cid}{w}}{\efalse} \mid \store \in \stores - \stores' \}\end{array}
$$

$$
\mathit{foldl}\ g\ \mems(S \cup F)\ \prog
$$

\begin{lemma}
  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
  for some $\cid$.
\end{lemma}

\begin{lemma}
  For all $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
  $$
  (\mathit{foldl}\ g\ \mems(S \cup F)\ \prog) =
  \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
  $$
\end{lemma}

\begin{lemma}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
  $$
  \pdf{S \cup F \cup V}(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
    (\mathit{foldl}\ g\ \mems(S \cup F)\ \prog) \\ 0 & \text{otherwise} \end{cases} 
  $$
  Then:
  $$
  \margd{\pdf{S \cup F \cup V}}{S \cup V} = \progd(\prog)
  $$
\end{lemma}

>>>>

\subsection{Method 1: Enumerating $\runs(\prog)$}

Here we describe a straightforward technique for obtaining program
distributions $\progd(\prog)$ through direct computation of
$\runs(\prog)$. An implementation of these techniques along with
Examples from Sections \ref{section-minicat-examples} and
\ref{section-metalang-ygc} is available online \cite{XXX}.

In Figure \ref{fig-solve} we define the algorithm $\solve{\stores}{\be}$
which filters $\stores$ based on satisfaction of $\be$. This algorithm
is fundamental to our brute force methods. It's correctness is characterized
as follows. 
\begin{lemma}
  For all $\stores$ and $\be$ with $\vars(\be) \subseteq \dom(\stores)$,
  $(\solve{\stores}{\be}) = \{ \store \in \stores \ \mid\ \lcod{\store,\be}{\cid} = 1 \}$
  for some $\cid$.
\end{lemma}
\begin{proof}
  Since we assume safety of programs we can assume that all variables in $\be$ have the
  same owner $\cid$. The result otherwise follows in a straightforward manner by induction
  on $\be$. 
\end{proof}

\begin{lemma}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{\etrue} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{\efalse} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
  \end{eqnarray*}
  Then $\cruns(\prog) = \runs(\prog)$.
\end{lemma}

\begin{lemma}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
  $$
  \pdf{S \cup F \cup V}(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
    \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
  $$
  Then:
  $$
  \margd{\pdf{S \cup F \cup V}}{S \cup V} = \progd(\prog)
  $$
\end{lemma}

\subsection{Method 2: Rewriting to Datalog}

### 2023-06-10 popl24/minicat.tex
checkpoint

Given a set of variables $X$, we write $\store_X$ to denoted the
memory $\store$ restricted to the domain $X$, and and we define
$\mems(X)$ as the set of all memories with domain $X$:
$$
\mems(X) \defeq \{ \store \mid \dom(\store) = X \}
$$
Thus, given a protocol $\prog$, the set of all random tapes for
$\prog$ is $\mems(\flips(\prog))$. We let $\stores$ range
over sets of memories.

@@

Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
any execution of $\prog$ is assumed to be an evaluation of a
configuration $\config{\store}{\prog}$ where $\store \in \mems(S \cup
F)$- that is, with secrets and the random tape as inputs- and where
$\config{\store}{\prog} \redxs \config{\store'}{\varnothing}$ it is
the case that $\store' \in \mems(S \cup F \cup V)$.

@@

\begin{example}[Oblivious Transfer]
  \label{example-lambda-OT}
  In \cite{XXX} a protocol for $\ttt{OT}$ is defined that we recreate as
  follows, where client 2 is the receiver and client 1 is the sender. Whereas
  in \cite{XXX} a trusted third party provides the sender and receiver with
  shared randomness, in our version both clients refer to the oracle $\ttt{H}$
  to obtain shared random values. 
\begin{verbatimtab}
  v[2, rd] := select(flip[2, d], H[r1], H[r0]);
  v[1, e] := s[2, c] xor flip[2, d];
  v[2, f0] := s[1, m0] xor select(v[1, e], H[r1], H[r0]);
  v[2, f1] := s[1, m1] xor select(not v[1, e], H[r1], H[r0]);
  v[2, mc] := select(s[2, c], v[2, f1], v[2, f0]) xor v[2, rd]
\end{verbatimtab}
Below we will discuss how to automatically verify security properties
of this protocol that correspond to the properties proven manually in
\cite{XXX}, and additionally how to automatically verify its passive
security in the simulator model. With respect to our language primitives,
we note that this protocol is probabilistically equivalent to the
following:
\begin{verbatimtab}  
  view[2, mc] := OT(s[2, c], s[1, m1], s[1, m0])
\end{verbatimtab}
\end{example}

>>>>

Given a set of variables $X$, we write $\store_X$ to denoted the
memory $\store$ restricted to the domain $X$, and and we define
$\mems(X)$ as the set of all memories with domain $X$:
$$
\mems(X) \defeq \{ \store \mid \dom(\store) = X \}
$$
Thus, given a protocol $\prog$, the set of all random tapes for
$\prog$ is $\mems(\flips(\prog))$. We let $\stores$ range
over sets of memories with the same domain, and abusing notation
we write $\dom(\stores)$ to denote the common domain.

@@

Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$,
any execution of $\prog$ is assumed to be an evaluation of a
configuration $\config{\store}{\prog}$ where $\store \in \mems(S \cup
F)$- that is, with secrets and the random tape as inputs- and where
$\config{\store}{\prog} \redxs \config{\store'}{\varnothing}$ it is
the case that $\store' \in \mems(S \cup F \cup V)$. We define a
program as \emph{safe} iff for all $\store \in \mems(S \cup F)$ there
exists $\store'$ where $\config{\store}{\prog} \redxs
\config{\store'}{\varnothing}$. Unsafe programs are those where views
are used before they're defined, or where expressions mix owners of
variables. As a sanity condition we will only consider safe programs
in our presentation, though in Section \ref{section-composition} we
define a type system that guarantees program safety.

@@

\begin{example}[Oblivious Transfer]
  \label{example-lambda-OT}
  In \cite{XXX} a protocol for $\ttt{OT}$ is defined that we recreate as
  follows, where client 2 is the receiver and client 1 is the sender. Whereas
  in \cite{XXX} a trusted third party provides the sender and receiver with
  shared randomness, in our version both clients refer to the oracle $\ttt{H}$
  to obtain shared random values. 
\begin{verbatimtab}
  v[2, rd] := select(flip[2, d], H[r1], H[r0]);
  v[1, e] := s[2, c] xor flip[2, d];
  v[2, f0] := s[1, m0] xor select(v[1, e], H[r1], H[r0]);
  v[2, f1] := s[1, m1] xor select(not v[1, e], H[r1], H[r0]);
  v[2, mc] := select(s[2, c], v[2, f1], v[2, f0]) xor v[2, rd]
\end{verbatimtab}
Below we will discuss how to automatically verify security properties
of this protocol that correspond to the properties proven manually in
\cite{XXX}, and additionally how to automatically verify its passive
security in the simulator model. With respect to our language primitives,
we note that this protocol is probabilistically equivalent to the
following:
\begin{verbatimtab}  
  v[2, mc] := OT(s[2, c], s[1, m1], s[1, m0])
\end{verbatimtab}
\end{example}
